---
title: "QSS Tidyverse Tutorial 0: Introduction to R"
tutorial:
  id: "00-intro"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will introduce you to the basics of R."
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
library(tidyverse)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tidyverse Tutorial 0: Output Report"
```


## Working with Data in tidyverse

### Subsetting a data frame (I)


You'll often need to access different parts of a data frame to use in other commands. For instance, maybe you want to take the mean of a column of the data frame or maybe you want to see all of the data for the 4th unit. Either way, we'll need to know how to subset the data frame. To select a particular variable from the data frame, you can use the `$` operator. So `mydata$myvar` will be a vector of just the `myvar` column of the `mydata` data frame.


#### Exercise

- Use the `$` to print out the `world.pop` variable from the `UNpop` data frame.

```{r dollar, exercise = TRUE, exercise.setup = "unpop"}
## print out the world.pop variable using $

```

```{r dollar-solution}
UNpop$world.pop
```

```{r dollar-check}
grade_code()
```


- Using the `$` operator, calculate the mean of the `world.pop` variable.

```{r meandollar, exercise = TRUE, exercise.setup = "unpop"}
## calculate the mean world population over this time period

```

```{r meandollar-solution}
mean(UNpop$world.pop)
```

```{r meandollar-check}
grade_code()
```

### Subsetting a data frame (II)

You can also use the brackets `[ ]` to subset from the data frame. But how will R know if you want to subset the rows or the columns? With a data frame as opposed to a vector, you will use a comma and the bracket will have the following form: `[rows, columns]` where the expression before the comma will select the rows and the expression after the comma will select the columns.

- `mydata[,"myvar"]` will select the `myvar` column from `mydata`
- `mydata[1,]` will select the first row of `mydata`
- `mydata[c(1,2,3),]` will select the first three rows of `mydata`
- `mydata[1:3, "myvar"]` will select the first three values of the `myvar` variable of `mydata`

#### Exercise


- Use the bracket commands to extract and print the `world.pop` variable from the `UNpop` data frame.

```{r brackets, exercise = TRUE, exercise.setup = "unpop"}
## use brackets to print out the world.pop variable

```

```{r brackets-solution}
UNpop[, "world.pop"]
```

```{r brackets-check}
grade_code()
```



- Use the bracket commands to extract and print rows 5 through 7 of the `UNpop` data frame.


```{r rows, exercise = TRUE, exercise.setup = "unpop"}
## extract rows 5 through 7 and all variables

```

```{r rows-solution}
UNpop[5:7, ]
```

```{r rows-check}
grade_code()
```


- Use the bracket commands to extract and print rows 5 through 7 of the `world.pop` variable of the `UNpop` data frame.


```{r rowcols, exercise = TRUE, exercise.setup = "unpop"}
## extract values 5 through 7 of the world.pop variable

```

```{r rowcols-solution}
UNpop[5:7, "world.pop"]
```

```{r rowcols-check}
grade_code()
```


## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
