---
title: "QSS Tutorial 6: Prediction II"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: >
  This tutorial will cover more advanced regression, including
  interactions and nonlinearities.
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tutorial 6: Output Report"
n <- 10
data(progresa, package = "qss")
```

## Extending Linear Regression



### Conceptual questions

```{r concept, echo=FALSE}
quiz(caption = "",
     question("How can regression models be used to draw causal inference?",
              answer("by predicting counterfactual outcomes", correct = TRUE),
              answer("by randomizing outcomes"),
              answer("by fitting a line")),
     question("Even if the average treatment effect is positive, the same treatment may not affect everyone in the positive direction. This is an example of:",
              answer("heterogenous treatment effects", correct = TRUE),
              answer("homogenous treatment effects"),
              answer("both of these"))
     )
```


### First regression

Today we will use a dataset designed to investigate the impact of Mexico's conditional cash transfer program, _Progresa_, on political outcomes such as voter turnout. See Section 4.5.2 in QSS for more information on this study. We have loaded the `progresa` data into this tutorial session. First, run a regression of turnout (`t2000`) on treatment status (`treatment`) to get a sense for the estimated effect of receiving the program earlier versus later. Save the regression to the object `fit` and print that object.

```{r progresa, exercise = TRUE}


```

```{r progresa-hint}
fit <- lm(_y_ ~ _x_, data = _dataname_)
fit
```

```{r progresa-check}
grade_result(
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment, data = progresa))),
          "Right! Remember that when you pass the `data` argument, you don't need to explicitly refer to the data frame in the formula."),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment))),
          "Good job! You might try using the `data` argument next time.")
)
```

### Add another covariate

Now, add the variable `avgpoverty` as an additional control variable to the regression (saving it as `fit_2var`) and print that new model.

```{r progresa-2var, exercise = TRUE}


```

```{r progresa-2var-solution}
fit_2var <- lm(t2000 ~ treatment + avgpoverty, data = progresa)
fit_2var
```

```{r progresa-2var-check}
grade_result(
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment + avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment + progresa$avgpoverty, data = progresa))),
          "Right! Remember that when you pass the `data` argument, you don't need to explicitly refer to the data frame in the formula."),
  pass_if(~ isTRUE(all.equal(.result, coef(lm(progresa$t2000 ~ progresa$treatment + progresa$avgpoverty)))),
          "Good job! You might try using the `data` argument next time.")
)
```


### Estimating interactions

Now, let's see how the effect of treatment varies by the average level of poverty. To do this, specify an interaction between `treatment` and `avgpoverty` using the `:` or `*` operators. Save this new model as `fit_int` and print it after running it.

```{r progresa-int, exercise = TRUE}

```

```{r progresa-int-solution}
fit_int <- lm(t2000 ~ treatment * avgpoverty, data = progresa)
fit_int
```

```{r progresa-int-check}
grade_result(
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment * avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment + avgpoverty + treatment * avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment + avgpoverty +  treatment:avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment * progresa$avgpoverty, data = progresa))),
          "Right! Remember that when you pass the `data` argument, you don't need to explicitly refer to the data frame in the formula."),
  pass_if(~ isTRUE(all.equal(.result, coef(lm(progresa$t2000 ~ progresa$treatment * progresa$avgpoverty)))),
          "Good job! You might try using the `data` argument next time.")
)
```


### Nonlinearities

Finally, let's investigate the relationship between poverty and turnout. Run a regression of turnout on `avgpoverty` and it squared term to account for any possible nonlinearities. Save the output as `fit_sq` and print it out.

```{r nonlin, exercise = TRUE}

```

```{r nonlin-solution}
fit_sq <- lm(t2000 ~ avgpoverty + I(avgpoverty ^ 2), data = progresa)
fit_sq
```

```{r nonlin-check}
grade_code()
```

## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
