---
title: "QSS Tutorial 8: Probability II"
output:
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
description: >
  This tutorial will cover random variables and their distributions.
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
learnr:::install_knitr_hooks()
#initialize_tutorial()
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tutorial 8: Output Report"
```

## Probability II

### Uniform distribution

What is the probability that a random variable with a uniform distribution in the range [1, 4] takes on a value less than or equal to 2.5?

```{r uniform, exercise = TRUE}

```

```{r uniform-check}
grade_result(pass_if(~ identical(.result, 0.5)))
```

### Variance

```{r var, echo = FALSE}
question("Suppose I give you a random variable X, such that its mean is 0 and that the expectation of X^2 is 5. What is the variance of X?",
         answer("2"),
         answer("3"),
         answer("4"),
         answer("5", correct = TRUE))
```

```{r var2, echo = FALSE}
question("If the variance of X is b, what is the variance of 2*X?",
         answer("0.5 * b"),
         answer("b"),
         answer("2 * b"),
         answer("4 * b", correct = TRUE))
```

### Voters

Suppose there are 10 voters in an election for two candidates A and B. Each voter independently chooses to vote for A with probability 0.1, and for B with probability 0.9. 

```{r voters, echo = FALSE}
question("What is the approximate probability that A gets at least 1 vote?",
         answer("0.65", correct = TRUE),
         answer("0.5"),
         answer("0.75"),
         answer("0.25"))
```

### Normal distribution

```{r normal, echo = FALSE}
question("Roughly what percentage of data with a standard normal distribution falls within +/- one standard deviation from the mean?",
         answer("33%"),
         answer("50%"),
         answer("66%", correct = TRUE),
         answer("87%"))
```


### Distribution of the mean

Suppose the standard deviation of a given population is 10. What is the standard deviation of sample mean within a random sample of size 25 from that population?

```{r clt, exercise = TRUE}

```

```{r clt-check}
grade_result(pass_if(~ .identical(.result, 2)))
```


## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
