---
title: "QSS Tutorial 8: Probability II"
output:
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
description: >
  This tutorial will cover random variables and their distributions.
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
learnr:::install_knitr_hooks()
#initialize_tutorial()
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tutorial 7: Output Report"
```

## Prediction 

### Uniform distribution

What is the probability that a random variable with a uniform distribution in the range [1, 4] takes on a value less than or equal to 2.5?

```{r uniform, exercise = TRUE}

```

```{r uniform-check}
grade_result(pass_if(~ identical(.result, 0.5)))
```

### Variance

```{r var, echo = FALSE}
question("Suppose I give you a random variable X, such that its mean is 0 and that the expectation of X^2 is 5. What is the variance of X?",
         answer("2"),
         answer("3"),
         answer("4"),
         answer("5", correct = TRUE))
```

```{r var2, echo = FALSE}
question("If the variance of X is b, what is the variance of 2*X?",
         answer("0.5 * b"),
         answer("b"),
         answer("2 * b"),
         answer("4 * b", correct = TRUE))
```

### Voters

Suppose there are 10 voters in an election for two candidates A and B. Each voter independently chooses to vote for A with probability 0.1, and for B with probability 0.9. 

```{r voters, echo = FALSE}
question("What is the approximate probability that A gets at least 1 vote?",
         answer("0.65", correct = TRUE),
         answer("0.5"),
         answer("0.75"),
         answer("0.25"))
```

### Normal distribution

```{r normal, echo = FALSE}
question("Roughly what percentage of data with a standard normal distribution falls within +/- one standard deviation from the mean?",
         answer("33%"),
         answer("50%"),
         answer("66%", correct = TRUE),
         answer("87%"))
```


### Central limit theorem

Suppose the standard deviation of a given population is 10. What is the standard deviation of sample mean within a sample of size 25 from that population?

```{r clt, exercise = TRUE}

```

```{r clt-check}
grade_result(pass_if(~ .identical(.result, 2)))
```

### Conceptual questions

```{r concept, echo=FALSE}
quiz(caption = "",
     question("How can regression models be used to draw causal inference?",
              answer("by predicting counterfactual outcomes", correct = TRUE),
              answer("by randomizing outcomes"),
              answer("by fitting a line")),
     question("Even if the average treatment effect is positive, the same treatment may not affect everyone in the positive direction. This is an example of:",
              answer("heterogenous treatment effects", correct = TRUE),
              answer("homogenous treatment effects"),
              answer("both of these"))
     )
```



### Bounding probabilities

Suppose that P(A)=0.4 and P(B)=0.3. What is the minimum value of P(A and not B)?

```{r ltp, exercise = TRUE}

```

```{r ltp-check}
grade_result(pass_if(~ identical(.result, 0.1)))
```

<div id="filter-hint">
**Hint:** See the Law of Total Probability. What happens if B is entirely contained in A?
</div>


### Combinations/Permutations

```{r combn, echo = FALSE}
question("Out of all possible anagrams for the word CHAIR (including the original word), what is the probability that the letter I comes first?",
         answer("0.1"),
         answer("0.2", correct = TRUE),
         answer("0.3"),
         answer("0.4"))
```


### Independence

Let A and B be two independent variables that take on only the values 0 and 1. If P(A=1, B=1)=0.12, P(A=0, B=1)=0.28, and P(A=1)=0.3, what is P(A=0, B=0)?

```{r indep, exercise = TRUE}

```

```{r indep-check}
grade_result(pass_if(~ identical(.result, 0.42)))
```


### Bayes' Rule

A disease affects 1% of the population. The only test for the disease returns positive 99% of the time the disease is actually there, though it returns positive 20% of the time if the disease is not actually there. You take this test and it returns positive. What is the probability you have the disease (rounded to nearest 0.01)?

```{r bayes, exercise = TRUE}

```


```{r exercise-check}
grade_result(pass_if(~ identical(.result, 0.05)))
```

### A couple of coins

You have a bucket with 2 coins: one is a regular coin, while the other has heads on both sides. You pick a coin at random and flip it: it turns up heads. What is the probability you chose the fair coin (express to the nearest 0.01)?

```{r coins, exercise = TRUE}

```

```{r coins-check}
grade_result(pass_if(~ .identical(.result, 0.33)))
```

## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
