---
title: "QSS Tidyverse Tutorial 3: Measurment I"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: >
  This tutorial will cover basic descriptive statistics and
  data visualization. Companion to Chapter 3 of "Quantitative Social
  Science: An Introduction in the Tidyverse."
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
library(tidyverse)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tidyverse Tutorial 3: Output Report"
data(STAR, package = "qss")
star <- STAR
```

## Handling Missing Data in R

### Small class size data


In this chapter, you'll analyze data from the STAR project, which is a four-year randomized trial on the effectiveness of small class sizes on education performance. The `star` data frame as been loaded into your space so that you can play around with it a bit.

Different from the baseR functions, we will teach you how to use tidyverse to perform the same analysis in this tutorial. 

#### Exercises
- The `Tibble` package is a core part tidyverse. It allows you to convert your traditional R data frames into tibbles, which are data frames too. But tibbles make it easier and faster to work with tidyverse. You can coerce a current data frame to a tibble with `as_tibble`.

```{r as_tibble-star, exercise = TRUE}

```

```{r as_tibble-star-solution}
as_tibble(star)
```

```{r as_tibble-star-check}
grade_code()
```

- Use the `str` function on the `star` to see what the data looks like. You can always use `head` function to view the first six rows of the data set.

```{r str-star, exercise = TRUE}

```

```{r str-star-solution}
str(star)
```

```{r str-star-check}
grade_code()
```


- Use the `glimpse` function on the `star` to see what the data looks like. `glimpse` is particularly useful when the data set contains a long list of variables (columns), as it allows you to see every column by transposing the original data set. 

```{r glimpse-star, exercise = TRUE}

```

```{r glimpse-star-solution}
print(star)
```

```{r glimpse-star-check}
grade_code()
```

- Use the `dim_desc` function on the `star` to see what the dimensions of the data look like.

```{r dim_desc-star, exercise = TRUE}

```

```{r dim_desc-star-solution}
dim_desc(star)
```

```{r dim_desc-star-check}
grade_code()
```


- Use the `summary` function on the `star` to get a sense for each variable.

```{r summary-star, exercise = TRUE}

```

```{r summary-star-solution}
summary(star)
```

```{r summary-star-check}
grade_code()
```



### Data wrangling with tidyverse: Handling missing data

From the baseR version of the qss tutorials, you have already learnt the basic toolkit that allows you to identify and work with the missing values (`NA`s) in your data set. 

Tidyverse also provides helpful tools for you to deal with the missing values. A value can be missing in one of two possible ways: a) explicitly as shown by a `NA` value; b) implicitly, as a value can simply not exist in the data. We use tidyverse to address the explicit missing values in this tutorial.  

#### Exercises
- Use the `drop_na` functions to drop the rows with explicit missing values in variable `g4math`. However, it is vital to make sure that you are not dropping the observations that are important in other representations of the data. So be careful when you use `drop_na`.

```{r drop_na, exercise = TRUE}

```

```{r drop_na-solution}
star %>% drop_na(g4math)
```

```{r drop_na-check}
grade_code()
```

- fill()
- replace_na()

## Visualizing Data

### Barplots


The **barplot** is a useful way to visualize a categorical or factor variable. In this exercise, you are going to visualize the `classtype` variable from the `star` data frame, which can take on the following values:

- `1` = small class
- `2` = regular class
- `3` = regular class with aid

#### Exercises
- Use the `group_by` and `count` function to create a data frame of counts `classcounts` for each category of the `classtype` in the `star` data.
- Use `geom_bar` function to plot a barplot for the `classcounts` data frame. Remeber to state `geom_bar(stat='identity')` to make sure ggpplot takes the correct input for x and y axes.
- You are encouraged to use pipes.

```{r barplot, exercise = TRUE}
## creat a data frame called classcounts that has
## the counts of each category of classtype


## make a barplot with ggplot

```

```{r barplot-solution}
classcounts <- star %>%
  group_by(classtype) %>%
  count() # of uniques

## make a barplot with ggplot
classcounts %>% ggplot(aes(x = classtype, y = n)) +
  geom_bar(stat='identity') 

```

```{r barplot-check}
grade_code("Awesome. The graph is looking a little unhelpful, though. Let's spruce it up.")
```

### Making the barplot readable

The default barplot usually isn't all that readable.

#### Exercises

- Use the function `scale_x_discrete(labels = c())` to name the categories of the dependent variable. Use `"Small class"` for 1, `"Regular class"` for 2, and `"Regular class with aid"` for 3. Remember to factorize the variable `classtype` in `aes()` with `factor()`, otherwise the x values will be defined as numeric rather than discrete. 
- You are encouraged to use pipes.

```{r barplot-labels, exercise = TRUE}
## Use the function scale_x_discrete(labels = c()) 
## i.e., scale_x_discrete(labels = c(
##     "1" = "Small class", 
##     "2" = "" ,
##     ...))



```

```{r barplot-labels-solution}
classcounts %>% ggplot(aes(x=factor(classtype), y = n)) +
  geom_bar(stat='identity') + 
   scale_x_discrete(labels = c(
     "1" = "Small class", 
     "2" = "Regular class", 
     "3" = "Regular class with aid"))

```


- Use the function `labs` to add the lables and title to the plot. Lable the x axis `"Classnames"`, the y axis `"Number of students"`, and the title `"The Distribution of Students in Different Class Types"`.
- You are encouraged to use pipes.


```{r barplot-lab, exercise = TRUE}
## Use the function `labs` and `title`



```

```{r barplot-lab-solution}
classcounts %>% ggplot(aes(x=factor(classtype), y = n)) +
  geom_bar(stat='identity') + 
   scale_x_discrete(labels = c(
     "1" = "Small class", 
     "2" = "Regular class", 
     "3" = "Regular class with aid")) +
  labs( x = "Classnames", 
        y = "Number of students", 
        title = "The Distribution of Students in Different Class Types")
```


### Histograms

For quantitative (numerical) variables, the barplot won't work because there are too many unique values. In this case, you will often use a histogram to visualize the a numerical variable.

#### Exercises
- With `ggplot()` function, use the `geom_histogram()` geom to create a histogram for the `g4math` variable in the `star` data frame.
- You are encouraged to use pipes.

```{r hist, exercise = TRUE}
## create a histogram of g4math with ggplot

```


```{r hist-solution}
## create a histogram of g4math with ggplot
star %>% ggplot(aes(x = g4math)) +
  geom_histogram()
```

```{r hist-check}
grade_code("Great job, though the graph is a bit spartan. Let's make it more readable.")
```

### Sprucing up the histogram

Same as the barplot, there are several arguments you can pass to the `ggplot()` function that will improve its readability:

- `aes(y = ..density..)` argument in `geom_histogram` allows you to make a density plot with `ggplot`.
- `labs` allows you to add character strings that prints a main title for the plot, and set the labels for the x (horizontal) and y (vertical) axes.
- `lims`: to specify the interval for the x and y axes.

#### Exercises
- Create a histogram with `ggplot` where you (a) include `aes(y = ..density..)` argument in `geom_histogram` to make a density plot, (b) set the y-axis to be between `0` and `0.015` using the `lims` argument, (c) include an informative x-axis label using the `labs` argument, and (d) include a title for the plot using the `labs` argument.
- Make sure to separate the arguments in function calls with commas.
- You are encouraged to use pipes.


```{r hist-nice, exercise = TRUE}
## create the histogram with the specifications given in the instructions

```


```{r hist-nice-solution}
## create the histogram with the specifications given in the instructions
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = ..density..)) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores")
```

### Adding lines and text to a plot

We'll often want to add more information to a plot to make it even more readable. You can do that with geoms that add to the current plot, such as `geom_abline` and `annotate`.

- `geom_abline`: adds a line with specific slope and intercept
- `geom_vline`: adds a vertical line
- `geom_hline`: adds a horizontal line

#### Exercises
- Use the `geom_vline` function to add a vertical line at the mean of the `g4math` variable from the `star` data. By default, missing values are removed by ggplot with a warning and you can use the `na.rm = TRUE` argument to silently remove them.
- You are encouraged to use pipes.

```{r hist-line, exercise = TRUE}
## add a vertical line at the mean of the variable
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = ..density..)) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores") +
  geom_vline(data = , aes(xintercept = mean()))
```

```{r hist-line-solution}
## add a vertical line at the mean of the variable with geom_vline
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = ..density..)) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores") +
  geom_vline(data = star, aes(xintercept = mean(g4math, na.rm = TRUE)))

```

```{r hist-line-check}
grade_code()
```

### Adding text to a plot

We'll sometimes want to add text to a plot to make it more informative. `annotate(geom = "text", x = 8, y = 9, label = "A")` adds a character string `A` centered at point on the (`8`, `9`) on the plot. You can use the axis labels to see where you might want to add these parts of the plot.

#### Exercise

- Use the `annotate` function to add the string `Average Score` to the plot at the point (750, 0.014).
- Make sure to separate the arguments in function calls with commas.
- You are encouraged to use pipes.

```{r hist-text, exercise = TRUE}
## add the text "Average Score" at the specified location

```

```{r hist-text-solution}
## add the text "Average Score" at the specified location
star %>% ggplot(aes(x = g4math)) +
  geom_histogram(aes(y = ..density..)) +
  lims(y = c(0, 0.015)) +
  labs(x = "Score",
       title = "Distribution of fourth-grade math scores") +
  annotate(geom = "text", x = 750, y = 0.014, label = "Average Score")
  

```

```{r hist-text-check}
grade_code()
```


## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
