---
title: "QSS Tidyverse Tutorial 4: Measurment II"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: >
  This tutorial will cover basic descriptive statistics and
  data visualization in tidyverse.
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
library(tidyverse)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tutorial 4: Output Report"
data(STAR, package = "qss")
star <- STAR
```

## Conceptual Questions


```{r concept, echo=FALSE}
quiz(caption = "",
     question("In section 3.6 of QSS, what kind of plot are we using to visualize ideological differences between Democrats and Republicans?",
              answer("scatterplot", correct = TRUE),
              answer("bar plot"),
              answer("histogram")),
     question("To estimate the correlation between two variables x and y, we need to know the mean of each as well as the:",
              answer("maximum"),
              answer("standard deviation", correct = TRUE),
              answer("length")),
     question("What type of relationships does correlation measure?",
              answer("linear relationships", correct = TRUE),
              answer("nonlinear relationships"),
              answer("both"))
     )
```



## More visualization


We'll continue to analyze data from the STAR project, which is a four-year randomized trial on the effectiveness of small class sizes on education performance. The `star` data frame as been loaded into your space so that you can play around with it a bit.

### Boxplots

Boxplots are useful tools to visualize how the distribution of a continuous variable changes across levels of a categorical variable. There are several ways to specify a boxplot, but the most basic way to construct with tidyverse is as follow:

    ggplot(star, aes(y = cont.var, x = cat.var)) +
      geom_boxplot()

Here in the `ggplot` function, the first argument is name of the data frame that contains the two variables of interest, followed by `aes` to specify them on the boxplot. Most importantly, we want to specify the function `geom_boxplot` to produce a boxplot.

#### Exercises

- First, let's factorize the `classtype` variable by using `mutate` and `as.factor` functions. Then, pass the expression to the `ggplot` function using the `%>%` operator. Include informative x-axis and y-axis labels with `labs` function. Finally, use the `scale_x-discrete` function to give the class types more intuitive names. 

```{r boxplot, exercise = TRUE}
## factorize the variable `classtype`


## create a box plot with the characteristics specified in the instructions


```


```{r boxplot-solution}
star %>% 
  mutate(classtype = as.factor(classtype)) %>%
  ggplot(aes(x = classtype, y = g4math)) + 
              geom_boxplot() +
              labs(y = "Math Scores", x = "Class Type") +
  scale_x_discrete(labels = c("Small class", "Regular class", "Regular class with aid"))
```

### Scatter plots

Now you'll think more about to measure bivariate relationships---that is, the relationship between two variables. The `ggplot` function takes the `data` argument followed by `aes` to specify the variable-of-interests, more importantly, specify `geom_point` to create a scatter plot as such: `ggplot(data, aes(x = variable, y = variable)) + geom_point()`. You will use this plot to explore the relationship between math and reading test scores in the `star` data.

#### Exercises
- Create a scatter plot using the `ggplot` and `geom_point` function with `g4math` on the x-axis and `g4reading` on the y-axis, both from the `star` data.
- Remember you can access a particular variable in a data frame using the `$` symbol.

```{r scatter, exercise = TRUE}
## produce a scatterplot of g4math on the x-axis and g4reading on the y-axis

```

```{r scatter-solution}
## produce a scatterplot of g4math on the x-axis and g4reading on the y-axis
ggplot(star, aes(x = g4math, y = g4reading)) +
  geom_point()
```

```{r scatter-check}
grade_code("OK, great plot! Let's make it a bit more polished.")
```


### Plotting two sets of points

Often we want to plot certain pints differently than others. For instance, maybe you want to see how the relationship between math and reading scores differs between students in small classes versus those not in small classes. To do this, we can subset the target classtypes with `filter` then specify colors of the different classtypes with `aes` within the `geom_point` function.

Before we do so, make sure to factorize the `classtype` variable for `ggplot` to create the graph.

#### Exercises

- Change the variable class of `classtype` to factor by using the `as.factor()` function
- Subset the data frame to only include Small class and Regular class for this scatter plot
- Pass the above expressions to the `ggplot` function using the `%>%` operator
- Differentiate data points from Small and Regular class by specifying `color` in the aes nested in `geom_point` 


```{r plot-points, exercise = TRUE}

```


```{r plot-points-solution}
## tidyverse
star %>% 
  mutate(classtype = as.factor(classtype)) %>%
  filter(classtype == c("1", "2")) %>%
  ggplot(aes(x = g4math, y = g4reading)) + geom_point(aes(color = classtype)) 
  
```

```{r plot-points-check}
grade_this_code()
```

### Finalizing your scatter plot

The scatter plot is looking very good, but it could use a little bit of polish. Let's add axis labels and a title. Remember that you can add these to the plot via the `xlab`, `ylab`, and `ggtitle`. Also, it's important to note that you can only add these to the initial `ggplot` function; adding them to the `points` or `text` or `abline` functions won't have any effect.

#### Exercises
- Take the current plot and add the following arguments: `xlab`, `ylab`, and `ggtitle`. Make them informative.

```{r niceplot, exercise = TRUE}

```


```{r niceplot-solution}
star %>% 
  mutate(classtype = as.factor(classtype)) %>%
  filter(classtype == c("1", "2")) %>%
  ggplot(aes(x = g4math, y = g4reading)) +
    geom_point(aes(color = classtype)) +
    labs(x = "Fourth Grade Math Scores", y = "Fourth Grade Reading Score", 
         title = "Math vs Reading")
```

## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
