---
title: "QSS Tutorial Week 5: Prediction I"
output:
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
description: >
  This tutorial will cover basic loops and if statements
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
learnr:::install_knitr_hooks()
#initialize_tutorial()
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tutorial Week 4: Output Report"
n <- 10
```

## Prediction 



### Conceptual questions

```{r concept, echo=FALSE}
quiz(caption = "",
     question("True or False: Loops can simplify our programming code when we want to repeatedly execute code chunks over and over again.",
              answer("True", correct = TRUE),
              answer("False")),
     question("To iterate meanst to: ",
              answer("perform repeatedly", correct = TRUE),
              answer("perform once"),
              answer("skip over"))
     )
```


### Loops

Suppose we assign `n <- 10` how many times will the loop `for (i in 1:n)` iterate?

```{r n_times, exercise = TRUE}

```

```{r n_times-check}
grade_result(
  pass_if(~ identical(.result, 10), "Good job!")
)
```


### if statement

```{r ifs, echo=FALSE}
question("In the conditional expression `if (Y) {expression1}`, `expression1` will be executed if `Y` is:",
         answer("`TRUE`", correct = TRUE),
         answer("`FALSE`"))
```



```{r for-if, echo=FALSE}
question("A conditional statement that is nested within a loop is executed _____ the loop.",
         answer("inside", correct = TRUE),
         answer("outside"))
```


### Loop practice

It is often helpful to create an empty vector to store the computational output from a loop. You can think of this as a bookshelf with spots for the result of each iteration of the loop. 

Create a vector called `output` that is filled with `NA` values and has length `n`.

```{r holder, exercise = TRUE}
output <-  
```

```{r holder-check}
grade_result(
  pass_if(~ identical(output, rep(NA, n)), "Good job!")
)
```

### Printing to the screen

In R, it's often very useful to print something to the screen and this is especially helpful when using loops. This can help us diagnose problems in the loop. 

Use the `cat()` function to put together a phrase and the object `n`. The final output should read `The object n has value 10`.

```{r cat, exercise = TRUE}

```

```{r cat-solution}
cat("The object n has value", n)
```

```{r cat-check}
grade_code()
```

### unique 

We've added the following 


```{r unique, exercise = TRUE}
fruit <- c("apple", "orange", "banana", "orange", "apple", "apple")

```

```{r unique-solution}
unique(fruit)
```

```{r unique-check}
grade_code()
```

### FizzBuzz

Here is a variant of the classic "FizzBuzz" coding question. Write a for loop that executes the code
     cat("The following number i is divisible by 3", i, "\n")
for numbers from 1 to 50 (inclusive) only for the numbers divisible by 3. Include brackets for the conditional.

```{r fizzbuzz, exercise = TRUE}

```


```{r fizzbuzz-solution}
for(i in 1:50){if (i %% 3 == 0) { cat("The following number i is divisible by 3", i, "\n") }}
```

```{r fizzbuzz-check}
grade_code()
```

### Adding a condition

Continuing from the previous FizzBuzz code you wrote, if the number is divisible by 5 (but not by 3), execute the code 
   cat("The following number i is divisible by 5", i, "\n")

```{r fizzbuzz2, exercise = TRUE}
for(i in 1:50){
  if (i %% 3 == 0) {
    cat("The following number i is divisible by 3", i, "\n")
  }
}
```

```{r fizzbuzz2-solution}
for(i in 1:50){
  if (i %% 3 == 0) {
    cat("The following number i is divisible by 3", i, "\n")
  } else if (i %% 5 == 0) {
    cat("The following number i is divisible by 5", i, "\n")
  }
}
```

```{r fizzbuzz2-check}
grade_code()
```

## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
