---
title: "QSS Tidyverse Tutorial 6: Prediction II"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: >
  This tutorial will cover more advanced regression, including
  interactions and nonlinearities.
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(qsslearnr)
library(tidyverse)
library(modelr)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
tut_reptitle <- "QSS Tutorial 6: Output Report"
n <- 10
data(progresa, package = "qss")
```

## Extending Linear Regression


### Conceptual questions

```{r concept, echo=FALSE}
quiz(caption = "",
     question("How can regression models be used to draw causal inference?",
              answer("by predicting counterfactual outcomes", correct = TRUE),
              answer("by randomizing outcomes"),
              answer("by fitting a line")),
     question("Even if the average treatment effect is positive, the same treatment may not affect everyone in the positive direction. This is an example of:",
              answer("heterogenous treatment effects", correct = TRUE),
              answer("homogenous treatment effects"),
              answer("both of these"))
     )
```


### First regression

Today we will use a data set designed to investigate the impact of Mexico's conditional cash transfer program, _Progresa_, on political outcomes such as voter turnout. See Section 4.5.2 in QSS for more information on this study. We have loaded the `progresa` data into this tutorial session. First, run a regression of turnout (`t2000`) on treatment status (`treatment`) to get a sense for the estimated effect of receiving the program earlier versus later. Save the regression to the object `fit` and print that object.

```{r progresa, exercise = TRUE}


```

```{r progresa-hint}
fit <- lm(_y_ ~ _x_, data = _dataname_)
fit
```

```{r progresa-check}
grade_result(
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment, data = progresa))),
          "Right! Remember that when you pass the `data` argument, you don't need to explicitly refer to the data frame in the formula."),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment))),
          "Good job! You might try using the `data` argument next time.")
)
```

### Coefficients

- A coefficient tells you the estimated effect of $x$ on $y$ in your fitted model. You can use the `coef` function to view the coefficient(s) and the intercept. 

```{r coef, exercise = TRUE}
fit <- lm(t2000 ~ treatment, data = progresa)
```

```{r coef-solution}
fit <- lm(t2000 ~ treatment, data = progresa)

coef(fit)
```

```{r coef-check}
grade_code()
```


### Add predictions and residuals to the original data set
- Besides the coefficients, predicted values and residuals also give us a sense of our fitted model. It is sometimes useful to add them to the original data set as new columns so plots can be made more easily. This can be done by the `add_predictions` and `add_residuals` functions. Let's experiment with `add_predictions` first, and then view the variables in `progresa` by the `glimpse` function.

```{r addpred, exercise = TRUE}
fit <- lm(t2000 ~ treatment, data = progresa)

progresa <- progresa %>%
  add_predictions(_model_) #Hint: pass your fitted model to the add_predictions function
# View the variables in progresa 

```

```{r addpred-solution}
fit <- lm(t2000 ~ treatment, data = progresa)

progresa <- progresa %>%
  add_predictions(fit) 

glimpse(progresa)
```

```{r addpred-check}
grade_code()
```

- Then do the same with the `add_residuals` function, and view with `glimpse`.

```{r addresid, exercise = TRUE}
fit <- lm(t2000 ~ treatment, data = progresa)

progresa <- progresa %>%

# View the variables in progresa 

```

```{r addresid-solution}
fit <- lm(t2000 ~ treatment, data = progresa)

progresa <- progresa %>%
  add_residuals(fit)

glimpse(progresa)
```

```{r addresid-check}
grade_code()
```


### Add another covariate

Now, add the variable `avgpoverty` as an additional control variable to the regression (saving it as `fit_2var`) and print that new model.

```{r progresa-2var, exercise = TRUE}


```

```{r progresa-2var-solution}
fit_2var <- lm(t2000 ~ treatment + avgpoverty, data = progresa)
fit_2var
```

```{r progresa-2var-check}
grade_result(
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment + avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment + progresa$avgpoverty, data = progresa))),
          "Right! Remember that when you pass the `data` argument, you don't need to explicitly refer to the data frame in the formula."),
  pass_if(~ isTRUE(all.equal(.result, coef(lm(progresa$t2000 ~ progresa$treatment + progresa$avgpoverty)))),
          "Good job! You might try using the `data` argument next time.")
)
```

### Estimating interactions

Now, let's see how the effect of treatment varies by the average level of poverty. To do this, specify an interaction between `treatment` and `avgpoverty` using the `:` or `*` operators. Save this new model as `fit_int` and print it after running it.

```{r progresa-int, exercise = TRUE}

```

```{r progresa-int-solution}
fit_int <- lm(t2000 ~ treatment * avgpoverty, data = progresa)
fit_int
```

```{r progresa-int-check}
grade_result(
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment + avgpoverty +  treatment:avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment * avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(t2000 ~ treatment + avgpoverty + treatment * avgpoverty, data = progresa)))),
  pass_if(~ isTRUE(all.equal(.result, lm(progresa$t2000 ~ progresa$treatment * progresa$avgpoverty, data = progresa))),
          "Right! Remember that when you pass the `data` argument, you don't need to explicitly refer to the data frame in the formula."),
  pass_if(~ isTRUE(all.equal(.result, coef(lm(progresa$t2000 ~ progresa$treatment * progresa$avgpoverty)))),
          "Good job! You might try using the `data` argument next time.")
)
```


### Nonlinearities

Finally, let's investigate the relationship between poverty and turnout. Run a regression of turnout on `avgpoverty` and it squared term to account for any possible nonlinearities. Save the output as `fit_sq` and print it out.

```{r nonlin, exercise = TRUE}

```

```{r nonlin-solution}
fit_sq <- lm(t2000 ~ avgpoverty + I(avgpoverty ^ 2), data = progresa)
fit_sq
```

```{r nonlin-check}
grade_code()
```

## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
